

### Task 2. (Sniatynska Olena)
#### 		In this task, we discuss a real-life process that is well modelled by a Poisson distribution. As you remember, a Poisson random variable describes occurrences of rare events, i.e., counts the number of successes in a large number of independent random experiments. One of the typical examples is the __radioactive decay__ process.
		
#### Consider a sample of radioactive element of mass $m$, which has a big _half-life period_ \(T\); it is vitally important to know the probability that during a one second period, the number of nuclei decays will not exceed some critical level \(k\). This probability can easily be estimated using the fact that, given the _activity_ ${\lambda}$ of the element (i.e., the probability that exactly one nucleus decays in one second) and the number $N$ of atoms in the sample, the random number of decays within a second is well modelled by Poisson distribution with parameter $\mu:=N\lambda$. Next, for the sample of mass $m$, the number of atoms is $N = \frac{m}{M} N_A$, where $N_A = 6 \times 10^{23}$ is the Avogadro constant, and $M$ is the molar (atomic) mass of the element. The activity of the element, $\lambda$, is $\log(2)/T$, where $T$ is measured in seconds. 

#### Assume that a medical laboratory receives $n$ samples of radioactive element ${{}^{137}}\mathtt{Cs}$ (used in radiotherapy) with half-life period $T = 30.1$ years and mass \(m = \mathtt{team\, id \,number} \times 10^{-6}\) g each. Denote by $X_1,X_2,\dots,X_n$ the __i.i.d.  r.v.__'s counting the number of decays in sample $i$ in one second. 

1.  Specify the parameter of the Poisson distribution of \(X_i\) (you'll need the atomic mass of _Cesium-137_)  
2.  Show that the distribution of the sample means of \(X_1,\dots,X_n\) gets very close to a normal one as $n$ becomes large and identify that normal distribution. To this end,
    +  simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;
    +  repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and form the empirical cumulative distribution function \(\hat  F_{\mathbf{s}}\) of $\mathbf{s}$;
    +  identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} \(F\) of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} \(\hat F_{\mathbf{s}}\) and plot both __c.d.f.__'s on one graph to visualize their proximity (use the proper scales!);
    +  calculate the maximal difference between the two \textbf{c.d.f.}'s;
    +  consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.   
3.  Calculate the largest possible value of $n$, for which the total number of decays in one second is less than $8 \times 10^8$ with probability at least $0.95$. To this end,  
    +  obtain the theoretical bound on \(n\) using Markov inequality, Chernoff bound and Central Limit Theorem, and compare the results;  
    +  simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sum $s=x_1 + \cdots +x_n$;
    +  repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of sums;
    +  calculate the number of elements of the sample which are less than critical value ($8 \times 10^8$) and calculate the empirical probability; comment whether it is close to the desired level \(0.95\)

```{r}
k <- 1e3
n <- 5

lambda <- log(2) / 936230400 #  with half-life period in sec

mass = 21*10**(-6)
N <- mass/137*6*10**23  # mass/ atomic mass of Cs* Avogadro constant
mu <- N * lambda
sigma <- sqrt(mu/n)
print(mu)
test <- rpois(1000, mu) #random density for poisson distribution with parameter mu
x <- seq(min(test), max(test)) 

plot(x, dpois(seq(min(test), max(test)), mu),main = "Poisson distribution")
plot(x, dnorm(x, mean = mu, sd = sqrt(mu)), main = "Normal distribution")


#Here on this graph you can see that the distribution of the sample means of X1,...,Xn gets very close to a normal one when n becomes large
```

#### We can now plot ecdf and cdf

```{r}
#### Next, calculate the parameters of the standard normal approximation
# P(mu)/n ~ P(mu, mu/n), because poisson r.v. ~ normal r.v.4

k <- 1e3
n <- 5

# Plot c.d.f and e.c.d.f

for (val in c(50,10,5)){
  
  sample_means <- colMeans(matrix(rpois(val*k, lambda = mu), nrow=val))
  xlims <- c(mu-3*sigma,mu+3*sigma)
  Fs <- ecdf(sample_means)

  plot(Fs, xlim = xlims, ylim = c(0,1),col = "blue",lwd = 1,
     main = "Comparison of ecdf and cdf")
  curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)

}

# Calculate the maximal difference between the two c.d.f.’s
dif <- max(abs(Fs(x) - pnorm(x, mean = mu, sd = sigma)))
print(dif)
```

```{r}
# Calculate the maximal difference between the two c.d.f.’s
sigma <-  sd(sample_means)

set.seed(10)
for (i in 1:100) {
  n <- i
  sample_mean_sum <- colSums(matrix(rpois(n * k, lambda = mu), nrow=n)); length(sample_mean_sum)
  corr <- length(sample_mean_sum[sample_mean_sum < 8e8]) / length(sample_mean_sum)
  if (corr < 0.5){
    n <- i
    break
  }
}

print(n)

```
Using imperative method we calculated that our upper bound is equal to 12, which means after this number our condition would not holds.
```{r}

num = (0.05*8e8)/mu

print(num)

```

Chernoff bound is calculeted with this formula:
$$
P(\frac{S_{n}-n\mu}{\sigma\sqrt{n}}> \frac{8e8-n\mu}{\sigma\sqrt{n}})\leq e^{-(\frac{8e8-n\mu}{4\sigma\sqrt(n)})^2}
$$
Lets find n from this equation:
$$ln(e^{-(\frac{8*10^{8}-68091468*n}{4*3569.051*\sqrt(n)})^2}) = ln(.05)$$
And our upper bound by Chernoff is equal to $$n ≈ 12.7747$$

Central Limit Theorem:
  Here is the formula: $$
\frac{S_{n}-n\mu}{\sigma\sqrt{n}} \sim N(0, 1)
$$

From this we have:
$$\Phi(\frac{8e8-68091468*n}{3569.051*\sqrt{n}})\geq .95$$
With Standart normal table here we have:
$$\frac{8e8-68091468*n}{3569.051*\sqrt{n}} = 1.65$$
And with CLT our n os equal to:
$$n ≈ 11.7486$$
As we can say Markov's inequality is the most imprecise and it's better to use Cernoff's bound or CLT.
  

